Denotational Semantics of Gradual Typing using Synthetic Guarded Domain Theory

### 2024-01-26 paper-new/appendix.tex
updates to appendix

%%%%%%%%%%%%%%%
% Composition %
%%%%%%%%%%%%%%%
\begin{lemma}
  Let $c : A \rel A'$ and $c' : A' \rel A''$ be value relations,
  and let $P_A, P_{A'}$ and $P_{A''}$ be monoids of perturbations.
  Given a push-pull structure $\piv_c$ for $c$ with respect to $P_A$ and $P_{A'}$, and
  and a push-pull structure $\piv_{c'}$ for $c'$ with respect to $P_{A'}$ and $P_{A''}$,
  we can define a push-pull structure $\piv_{c \comp c'}$ for $c \comp c'$.

@@

%%%%%%%%%%%%%
% UF and FU %
%%%%%%%%%%%%%
\begin{lemma}
  Let $c : A \rel A'$ and let $P_A$ and $P_{A'}$ be monoids of perturbations.
  Given a push-pull structure $\piv_c$ for $c$ with respect to $P_A$ and $P_{A'}$,
  we can define a push-pull structure $\pie_{UF(c)}$ for $UF(c)$ with respect to $UF(P_{A})$
  and $UF(P_{A'})$.
  % Eric: Are we assuming here that F and U preserve perturbations?
\end{lemma}
\begin{proof}
  Follows by the functorial action of $UF$ and $FU$ on perturations and squares.
\end{proof}

%%%%%%%%%%%%
% Products %
%%%%%%%%%%%%
\begin{lemma}
  Let $c_1 : A_1 \rel A_1'$ and $c_2 : A_2 \rel A_2'$, and let $P_{A_1}$,
  $P_{A_2}$, $P_{A_1'}$ and $P_{A_2'}$ be monoids of perturbations.
  Given push-pull structures 
  $\piv_{c_1}$ with respect to $P_{A_1}$ and $P_{A_1'}$, and 
  $\piv_{c_2}$ with respect to $P_{A_2}$ and $P_{A_2'}$,
  we can define a push-pull structure $\piv_{c_1 \times c_2}$
  with respect to $P_{A_1} \times P_{A_2}$ and $P_{A_1'} \times P_{A_2'}$.

  Likewise, given push-pull structures 
  $\pie_{Fc_1}$ with respect to $F(P_{A_1})$ and $F(P_{A_1'})$, and
  $\pie_{Fc_2}$ with respect to $F(P_{A_2})$ and $F(P_{A_2'})$,
  we can define a push-pull structure $\pie_{F(c_1 \times c_2)}$
  with respect to $F(P_{A_1} \times P_{A_2})$ and $F(P_{A_1'} \times P_{A_2'})$.
\end{lemma}
\begin{proof}
We define the push function for $\piv_{c_1 \times c_2}$ by 
$\push(\delta_1, \delta_2) = (\push_{c_1}(\delta_1), \push_{c_2}(\delta_2))$ and likewise for $\pull$.
The push-pull property holds because it holds for each component.
\end{proof}

%%%%%%%%%
% Arrow %
%%%%%%%%%
\begin{lemma}
  Let $c : A \rel A'$ and $d : B \rel B'$.
  Given push-pull structures $\piv_c$ for $c$ and $\pie_d$ for $d$,
  we can define a push-pull structure $\pie_{c \arr d}$ for $c \arr d$.

  Given push-pull structures $\pie_{Fc}$ for $Fc$ and $\piv_{Ud}$ for $Ud$,
  we can define a push-pull structure $\piv_{U(c \arr d)}$ for $U(c \arr d)$
  with respect to the monoids $(P_{FA})^{op} \oplus P_B$
\end{lemma}
\begin{proof}
  The former follows from the functorial action of $\arr$ on perturbations and squares,
  and the latter uses the functorial action of $\tok$ on perturbations and squares.
\end{proof}

%%%%%%%%%%%%%%%%%%%%%%
% Model Construction %
%%%%%%%%%%%%%%%%%%%%%%
We now proceed with the construction of the model:
    % Write 
    % %
    % \[ \mathcal M = (\vf, \vsq, \ef, \esq, \Ff, \Fsq, \Uf, \Usq, \arrf, \arrsq). \] 
    % %

    Define a step-2 model $\mathcal M'$ as follows:
    \begin{itemize}
      \item Value objects are triples consisting of:
      \begin{itemize}
        \item A value object $A$ in $\vf$ 
        \item A monoid $P_A$ of ``pure'' perturbations that is a submonoid of 
        $\{ f \in \vf(A, A) \mid f \bisim \id_A \}$
        \item A monoid $P^K_A$ of ``impure'' or ``Kleisli'' perturbations that is a submonoid of
        $\{ \phi \in \ef(FA, FA) \mid \phi \bisim \id_{FA} \}$ such that $P^K_A$ contains
        $\delta_A^*$
      \end{itemize}

      \item Computation objects are triples consisting of:
      \begin{itemize}
        \item A computation object $B$ in $\ef$
        \item A monoid $P_B$ of ``pure'' perturbations that is a submonoid of
        $\{ \phi \in \ef(B, B) \mid \phi \bisim \id_B \}$
        \item A monoid $P^K_B$ of ``impure'' or ``Kleisli'' perturbations that is a submonoid
        of $\{ g \in \vf(UB, UB) \mid g \bisim \id_{UB} \}$
      \end{itemize}

@@

    % Before introducing the relations, we make a definition.

    % \begin{definition}[push-pull structure]
    %   Let $c : A \rel A'$ be a value relation of $\mathcal M$. A \emph{value push-pull structure} $\piv_c$ for $c$ consists of:
    %   \begin{itemize}
    %     \item A function $\push : P_A \to P_{A'}$ 
    %           such that for all $\delta^l \in P_A$ we have $\delta^l \ltdyn_c^c \push(\delta^l)$.
    %     \item A function $\push^K : P^K_A \to P^K_{A'}$ 
    %           such that for all $\delta^K_l \in P^K_A$ we have $\delta^K_l \ltdyn_{Fc}^{Fc} \push(\delta^K_l)$.
    %     \item A function $\pull : P_{A'} \to P_A$
    %           such that for all $\delta_r \in P_{A'}$ we have $\pull(\delta^r) \ltdyn_{c}^c \delta^r$.
    %     \item A function $\pull^K : P^K_{A'} \to P^K_A$
    %           such that for all $\delta^K_r \in P^K_{A'}$ we have $\pull(\delta^K_r) \ltdyn_{Fc}^{Fc} \delta^K_r$.
    %   \end{itemize}

    %   For $d : B \rel B'$ a computation relation, we define a \emph{computation push-pull structure} $\pie_d$ for $d$
    %   in an analogous manner.
    % \end{definition}

    % Now we continue with the description of the construction:
    \begin{itemize}

      \item The objects of $\vsq'$ (i.e., the value relations) are triples consisting of:
      \begin{itemize}
        \item A value relation $c \in \vsq$.
        \item A push-pull structure $\piv_c$ for $c$ with respect to $P_A$ and $P_{A'}$.
        \item A push-pull structure $\pie_{Fc}$ for $Fc$ with respect to $P^K_A$ and $P^K_{A'}$.
      \end{itemize}

      The objects of $\esq'$ are defined analogously, i.e.,
      \begin{itemize}
        \item A computation relation $d \in \esq$.
        \item A push-pull structure $\pie_d$ for $d$ with respect to $P_B$ and $P_{B'}$.
        \item A push-pull structure $\piv_{Ud}$ for $Ud$ with respect to $P^K_B$ and $P^K_{B'}$.
      \end{itemize}

@@

      \item We define composition of relations $(c, \piv_c, \pie_{Fc})$ and 
      $(c', \piv_{c'}, \pie_{Fc'})$ as $(c \comp c', \piv_{c \comp c'}, \pie_{Fc \comp Fc'})$
      and likewise for computation relations.

@@

      % \item We begin with the functor $m$ that composes relations and squares horizontally.
      % Given value relations $(c ,\piv_c, \pie_{Fc})$ and $(c', \piv_{c'}, \pie_{Fc'})$ where $c : A \rel A'$ and $c' : A' \rel A''$,
      % we define their composition to be the triple $(c \comp c', \piv_{c \comp c'})$ where we define
      % $\piv_{c \comp c'}$ as the following push-pull structure:

      % \begin{itemize}
      %   \item $\push_{c \comp c'} = \push_{c'} \circ \push_{c}$
      %   \item $\push^K_{c \comp c'} = \push^K_{c'} \circ \push^K_{c}$
      %   \item $\pull_{c \comp c'} = \pull_{c} \circ \pull_{c'}$
      %   \item $\pull^K_{c \comp c'} = \pull^K_{c} \circ \pull^K_{c'}$
      % \end{itemize}

      % We observe that each of the components satisfies the commuting necessary square.

@@

      \[ (A_1, P_{A_1}, P^K_{A_1}) \times (A_2, P_{A_2}, P^K_{A_2}) =
        (A_1 \times A_2, P_{A_1} \times P_{A_2}, P^K_{A_1} \oplus P^K{A_2}). \]

@@

      We define $\times$ on relations by

      \[ (c_1, \piv_{c_1}, \pie_{Fc_1}), (c_2 \piv_{c_2}, \pie_{Fc_2}) = 
        (c_1 \times c_2, \piv_{c_1 \times c_2}, \pie_{F(c_1 \times c_2)}). \]

@@

      \[ F(A, P_A, P^K_A) = (FA, P^K_A, UF(P_A)). \]

      We define $F$ on relations by

      \[ F(c, \piv_c, \pie_{Fc}) = (Fc, \pie_{Fc}, \piv_{UF(c)}). \]

@@

      \[ U(B, P_B, P^K_B) = (UB, P^K_B, FU(P_B). )\]

      We define $U$ on relations by

      \[ U(d, \pie_d, \piv_{Ud}) = (Ud, \piv_{Ud}, \pie_{FU(d)})  \]

@@

      \[ (A, P_A, P^K_A) \arr (B, P_B, P^K_B) = 
        (A \arr B, (P_A)^{op} \times P_B, (P^K_A)^{op} \oplus P^K_B). \]

      We define $\arr$ on relations by

      \[ (c, \piv_c, \pie_{Fc}) \arr (d, \pie_d, \piv_{Ud}) =
        (c \arr d, \pie_{c \arr d}, \piv_{U(c \arr d)}). \]

      % \item We define $F$ on objects by $F (A, \pv_A, \ptbv_A) = (FA, (1 + \pv_A), h_F)$
      % where $1$ is the trivial monoid, $+$ is the coproduct in the category of monoids, and $h_F$ is the homomorphism defined as follows:

      % \item We define $U$ on objects by $U (B, \pe_B, \ptbe_B) = (UB, \pe_B, h_U)$
      % where $h_U(p_B) = U(\ptbe_B(p_B))$.

      % \item We define $(A, \pv_A, \ptbv_A) \arr (B, \pe_B, \ptbe_B) = (A \arr B, \pv_A \times \pe_B, h_\arr)$
      % where $\times$ is the product in the category of monoids, and $h_\arr$ is defined by 
      % $h_\arr(p_A, p_B) = \ptbv_A(p_A) \arr \ptbe_B(p_B)$.
    \end{itemize}

@@

\begin{definition}[representation structure]
Let $c : A \rel A'$ be a value relation. A \emph{left-representation structure}
$\rho^L_c$ for $c$ consists of a value morphism $e_c \in \vf(A, A')$ such that
$c$ is quasi-left-representable by $e_c$ (see Definition \ref{def:quasi-left-representable}).

Likewise, let $d : B \rel B'$ be a computation relation. A \emph{right-representation structure}
$\rho^R_d$ for $d$ consists of a computation morphism $p_d \in \ef(B', B)$
such that $d$ is quasi-right-representable by $p_d$ (see Definition
\ref{def:quasi-right-representable}).
\end{definition}

(Notice that the direction of the morphism is opposite in the definition of
right-representation structure.)

@@

@@

    % 3.
    \item We define $\rho^R_{F(c \comp c')}$ as follows:
    \begin{itemize}
      \item $p_{F(c \comp c')} = p_{Fc} \circ p_{Fc'}$ % : FA'' \to FA
      \item $\dellp_{F(c \comp c')} = $
      \item $\delrp_{F(c \comp c')} = $
    \end{itemize}

    % 4.
    \item We define $\rho^L{(U(d \comp d'))}$ as follows:

>>>>

%%%%%%%%%%%%%%%
% Composition %
%%%%%%%%%%%%%%%
\begin{lemma}\label{lem:push-pull-comp}
  Let $c : A \rel A'$ and $c' : A' \rel A''$ be value relations,
  and let $P_A, P_{A'}$ and $P_{A''}$ be monoids of perturbations.
  Given a push-pull structure $\piv_c$ for $c$ with respect to $P_A$ and $P_{A'}$, and
  and a push-pull structure $\piv_{c'}$ for $c'$ with respect to $P_{A'}$ and $P_{A''}$,
  we can define a push-pull structure $\piv_{c \comp c'}$ for $c \comp c'$.

@@

%%%%%%%%%%%
% U and F %
%%%%%%%%%%%
\begin{lemma}\label{lem:push-pull-U-F}
  Let $c : A \rel A'$ and let $P_A$ and $P_{A'}$ be monoids of perturbations.
  Given a push-pull structure $\piv_c$ for $c$ with respect to $P_A$ and $P_{A'}$,
  we can define a push-pull structure $\pie_{Fc}$ for $F(c)$ with respect to
  $\mathbb{N} \times P_{A}$ and $\mathbb{N} \times P_{A'}$.

  Likewise, from a push-pull structure for $d$ with respect to $P_B$ and $P_{B'}$,
  we can define a push-pull structure for $Ud$ with respect to $\mathbb{N} \times P_B$
  and $\mathbb{N} \times P_{B'}$.
\end{lemma}
\begin{proof}

  We define $\push_{Fc} : \mathbb{N} \times P_{A} \to \mathbb{N} \times P_{A'}$.
  Since the product of commutative monoids coincides with the coproduct, it
  suffices by the universal property of the coproduct to consider the two cases separately.

  Given $n \in \mathbb{N}$ we define the push to be $n$.
  Then the required square is $\delta_{FA}^* \ltdyn_{Fc}^{Fc} \delta_{FA'}^*$ which exists by
  the definition of a step-1 model.
  On the other hand, given $a \in P_{A}$, we define the push to be $\push_c(a)$
  and then the required exists by the push property for $c$.
  Defining $\pull_{Fc}$ is done similarly.

  The proof for $Ud$ is analogous.
\end{proof}

%%%%%%%%%%%%
% Products %
%%%%%%%%%%%%
\begin{lemma}\label{lem:push-pull-times}
  Let $c_1 : A_1 \rel A_1'$ and $c_2 : A_2 \rel A_2'$, and let $P_{A_1}$,
  $P_{A_2}$, $P_{A_1'}$ and $P_{A_2'}$ be monoids of perturbations.
  Given push-pull structures 
  $\piv_{c_1}$ with respect to $P_{A_1}$ and $P_{A_1'}$, and 
  $\piv_{c_2}$ with respect to $P_{A_2}$ and $P_{A_2'}$,
  we can define a push-pull structure $\piv_{c_1 \times c_2}$
  with respect to $P_{A_1} \times P_{A_2}$ and $P_{A_1'} \times P_{A_2'}$.

  % Likewise, given push-pull structures 
  % $\pie_{Fc_1}$ with respect to $F(P_{A_1})$ and $F(P_{A_1'})$, and
  % $\pie_{Fc_2}$ with respect to $F(P_{A_2})$ and $F(P_{A_2'})$,
  % we can define a push-pull structure $\pie_{F(c_1 \times c_2)}$
  % with respect to $F(P_{A_1} \times P_{A_2})$ and $F(P_{A_1'} \times P_{A_2'})$.
\end{lemma}
\begin{proof}
We define the push function for $\piv_{c_1 \times c_2}$ by 
$\push(\delta_1, \delta_2) = (\push_{c_1}(\delta_1), \push_{c_2}(\delta_2))$ and likewise for $\pull$.
The push-pull property holds because it holds for each component.
\end{proof}

%%%%%%%%%
% Arrow %
%%%%%%%%%
\begin{lemma}\label{lem:push-pull-arrow}
  Let $c : A \rel A'$ and $d : B \rel B'$.
  Given push-pull structures $\piv_c$ for $c$ and $\pie_d$ for $d$,
  we can define a push-pull structure $\pie_{c \arr d}$ for $c \arr d$.

  % Given push-pull structures $\pie_{Fc}$ for $Fc$ and $\piv_{Ud}$ for $Ud$,
  % we can define a push-pull structure $\piv_{U(c \arr d)}$ for $U(c \arr d)$
  % with respect to the monoids $(P_{FA})^{op} \oplus P_B$
\end{lemma}
\begin{proof}
  Follows from the functorial action of $\arr$ on perturbations and squares.
  % and the latter uses the functorial action of $\tok$ on perturbations and squares.
\end{proof}

%%%%%%%%%%%%%%%%%%%%%%
% Model Construction %
%%%%%%%%%%%%%%%%%%%%%%
We now proceed with the construction of the model:

    Define a step-2 model $\mathcal M'$ as follows:
    \begin{itemize}
      \item Value objects are pairs consisting of:
      \begin{itemize}
        \item A value object $A$ in $\vf$.
        \item A commutative monoid $P_A$ of perturbations
        and a monoid homomorphism $\ptb_A : P_A \to \{ f \in \vf(A, A) \mid f \bisim \id_A \}$.

      \item Computation objects are pairs consisting of:
      \begin{itemize}
        \item A computation object $B$ in $\ef$.
        \item A commutative monoid $P_B$ of perturbations and a monoid homomorphism
        $\ptb_B : P_B \to \{ \phi \in \ef(B, B) \mid \phi \bisim \id_B \}$.
      \end{itemize}

@@

      \item The objects of $\vsq'$ (i.e., the value relations) are pairs consisting of:
      \begin{itemize}
        \item A value relation $c \in \vsq$.
        \item A push-pull structure $\piv_c$ for $c$ with respect to $P_A$ and $P_{A'}$.
        %\item A push-pull structure $\pie_{Fc}$ for $Fc$ with respect to $P^K_A$ and $P^K_{A'}$.
      \end{itemize}

      The objects of $\esq'$ are defined analogously, i.e.,
      \begin{itemize}
        \item A computation relation $d \in \esq$.
        \item A push-pull structure $\pie_d$ for $d$ with respect to $P_B$ and $P_{B'}$.
        %\item A push-pull structure $\piv_{Ud}$ for $Ud$ with respect to $P^K_B$ and $P^K_{B'}$.
      \end{itemize}

@@

      \item We define composition of relations $(c, \piv_c)$ and $(c', \piv_{c'})$
      as $(c \comp c', \piv_{c \comp c'})$ where $\piv_{c \comp c'}$ is as in Lemma
      \ref{lem:push-pull-comp}, and likewise for computation relations.

@@

@@

      \[ (A_1, P_{A_1}) \times (A_2, P_{A_2}) =
        (A_1 \times A_2, P_{A_1} \times P_{A_2}). \]

@@

      Using Lemma \ref{lem:push-pull-times}, we define $\times$ on relations by

      \[ (c_1, \piv_{c_1}), (c_2 \piv_{c_2}) = 
        (c_1 \times c_2, \piv_{c_1 \times c_2}). \]

@@

      \[ F(A, P_A, \ptb_A) = (FA, \mathbb{N} \times P_A, \ptb_{FA}), \]

      where we define $\ptb_{FA}(n, a) = (\delta_{FA}^*)^n \circ F(\ptb_A(a))$.

      Using Lemma \ref{lem:push-pull-U-F}, we define $F$ on relations by

      \[ F(c, \piv_c) = (Fc, \pie_{Fc}). \]

@@

      \[ U(B, P_B, \ptb_B) = (UB, \mathbb{N} \times P_B, \ptb_{UB}), \]

      where we define $\ptb_{UB}(n, b) = (\delta_{UB}^*)^n \circ U(\ptb_B(b))$.

      Using Lemma \ref{lem:push-pull-U-F}, we define $U$ on relations by

      \[ U(d, \pie_d) = (Ud, \piv_{Ud}).  \]

@@

      \[ (A, P_A, \ptb_A) \arr (B, P_B, \ptb_B) = 
        (A \arr B, P_A \times P_B, \ptb_{A \arr B}), \]

      where we define $\ptb_{A \arr B}(a, b) = \ptb_A(a) \arr \ptb_B(b)$.

      Using Lemma \ref{lem:push-pull-arrow}, we define $\arr$ on relations by

      \[ (c, \piv_c) \arr (d, \pie_d) =
        (c \arr d, \pie_{c \arr d}). \]

    \eric{Make sure this part is correct}
    We define a ``Kleisli arrow'' operation taking a perturbation in $P_{UB}$
    to a perturbation in $P_{U(A \to B)}$, as follows:
    Given a perturbation $u = (n, b) \in P_{UB}$ we define $\id \tok u$ to be 
    $(n, \id_{P_A}, b) \in P_{U(A \to B)}$.
    We need to show that
%
    \[ \ptb_{U(A \arr B)}(\id \tok (n, b)) = \id \tok \ptb_{UB}(n, b). \]
%
    Since the product and coproduct coincide in the category of commutative monoids,
    and since $\mathbb{N}$ is generated by the element $1$,
    it suffices by the universal property of the coproduct to show
%
    \[ \ptb_{U(A \arr B)}(\id \tok (1, \id_{P_B})) = \id \tok \ptb_{UB}(1, \id_{P_B}), \]
%
    and
%
    \[ \ptb_{U(A \arr B)}(\id \tok (0, b)) = \id \tok \ptb_{UB}(0, b). \]

    For the former, the LHS is
%
    \begin{align*}
      \ptb_{U(A \arr B)}&(\id \tok (1, \id_{P_B})) \\ 
      &= \ptb_{U(A \arr B)}(1, \id_{P_A}, \id_{P_B}) \\ 
      &= \delta_{U(A \arr B)}^* \circ U(\ptb_A(\id_{P_A}) \arr \ptb_B(\id_{P_B})) \\
      %&= \delta_{U(A \arr B)}^* \circ U(\id_A \arr \id_B) \\
      &= \delta_{U(A \arr B)}^*
    \end{align*}
%
    and we need to show that this is equal to 
    $\id \tok \ptb_{UB}(1, \id_{P_B}) = \id \tok \delta_{UB}^*$.
    But this holds because $\delta_{UB}^*$ commutes with all computation morphisms,
    in particular, with bind and application.

    For the latter, we need to show
%
    \[ U(\id \arr \ptb_B(b)) = \id \tok U(\ptb_B(b)), \]
%
    which holds by definition of $\tok$.

    Given a perturbation $q = (n, a) \in P_{FA}$ we define
    $(n, a) \tok \id = (n, a, \id_{P_B}) \in P_{U(A \to B)}$.
    We need to show that
  %
    \[ \ptb_{U(A \arr B)}((n, a) \tok \id) = \ptb_{FA}(n, a) \tok \id. \]
  %
    By similar reasoning to the above, it will suffice to show that

    \[ \delta_{U(A \arr B)}^* = \delta_{FA}^* \tok \id \]

    And this follows from the fact that $\delta_{FA}^*$ commutes with all computation morphisms.

    We can similarly construct the Kleisli product operation on perturbations.
    Given a perturbation $(n, a_1, a_2) \in P_{F(A_1 \times A_2)}$ we define
    $(n, a_1, a_2) \timesk \id = (n, a_1, \id_{P_{A_2}})$,
    and likewise we define $\id \timesk (n, a_1, a_2) = (n, \id_{P_{A_1}}, a_2)$.

% We have $\ptb_{U(A \arr B)}(\id \tok (n, b)) = \ptb_{U(A \arr B)}(n, \id, b)$, 
% which is equal to $\delta_{U(A \arr B)}^n (\id \to \ptb_B(b))$

@@

\begin{definition}[representation structure]
  Let $c : A \rel A'$ be a value relation. A \emph{left-representation structure}
  $\rho^L_c$ for $c$ consists of a value morphism $e_c \in \vf(A, A')$ such that
  $c$ is quasi-left-representable by $e_c$ (see Definition \ref{def:quasi-left-representable}).
  We similarly define a {right-representation structure} $\rho^R_c$ for $c$ to consist of a morphism
  $p_c \in \vf{A', A}$ such that $c$ is quasi-right-representable by $p_c$ (see Definition \ref{def:quasi-right-representable}).
  % (Notice that the direction of the morphism is opposite in the definition of
  % right-representation structure.)

  Likewise, let $d : B \rel B'$ be a computation relation. Left- and right-representation
  structures for $d$ are defined in a similar manner, except the representing morphisms
  are now computation morphisms rather than value morphisms.

  % Likewise, let $d : B \rel B'$ be a computation relation. A \emph{right-representation structure}
  % $\rho^R_d$ for $d$ consists of a computation morphism $p_d \in \ef(B', B)$
  % such that $d$ is quasi-right-representable by $p_d$ (see Definition
  % \ref{def:quasi-right-representable}).
\end{definition}

@@

\begin{definition}
  Let $c, c' : A \rel A'$. We say that $c$ and $c'$ are \emph{quasi-order-equivalent},
  written $c \qordeq c'$, if there exist perturbations $\delta^l_1, \delta^l_2 \in \pv_A$ and 
  $\delta^r_1, \delta^r_2 \in \pv_{A'}$ such that the following two squares exist:

  \begin{center}
    \begin{tabular}{ m{9em} m{9em} } 
      \begin{tikzcd}[ampersand replacement=\&]
        A \& {A'} \\
        A \& {A'}
        \arrow["\delta^l_1"', from=1-1, to=2-1]
        \arrow["\delta^r_1", from=1-2, to=2-2]
        \arrow["c", "\shortmid"{marking}, no head, from=1-1, to=1-2]
        \arrow["c'"', "\shortmid"{marking}, no head, from=2-1, to=2-2]
      \end{tikzcd}
      &
      \begin{tikzcd}[ampersand replacement=\&]
        A \& {A'} \\
        A \& {A'}
        \arrow["\delta^l_2"', from=1-1, to=2-1]
        \arrow["\delta^r_2", from=1-2, to=2-2]
        \arrow["c'", "\shortmid"{marking}, no head, from=1-1, to=1-2]
        \arrow["c"', "\shortmid"{marking}, no head, from=2-1, to=2-2]
      \end{tikzcd}
    \end{tabular}
  \end{center}

  We make the analogous definition for computation relations $d, d' : B \rel B'$.
\end{definition}

\begin{lemma}\label{lem:left-rep-by-same-morphism}
  Let $c, c' : A \rel A'$. If $c$ and $c'$ are both quasi-left-representable
  by the same $f$, then $c \qordeq c'$.
\end{lemma}
\begin{proof}
  By $\upr$ for $c'$, there exists a perturbation $\delle_{c'}$ and a square
  $\upr_{c'} : \delle_{c'} \ltdyn_{c'}^{r(A)} f$.

  By $\upl$ for $c$, there exists a perturbation $\delre_c$ and a square 
  $\upl_c : f \ltdyn_{r(A')}^{c} \delre_c$.

  Composing these horizontally we get the following square:

  % https://q.uiver.app/#q=WzAsNixbMCwwLCJBIl0sWzEsMCwiQSJdLFsyLDAsIkEnIl0sWzAsMSwiQSJdLFsxLDEsIkEnIl0sWzIsMSwiQSciXSxbMCwxLCJyKEEpIl0sWzEsMiwiYyJdLFszLDQsImMnIiwyXSxbNCw1LCJyKEEnKSIsMl0sWzAsMywiXFxkZWxsZV97Yyd9IiwyXSxbMSw0LCJmIiwyXSxbMiw1LCJcXGRlbHJlX2MiXV0=
  \[\begin{tikzcd}[ampersand replacement=\&]
    A \& A \& {A'} \\
    A \& {A'} \& {A'}
    \arrow["{r(A)}", from=1-1, to=1-2]
    \arrow["c", from=1-2, to=1-3]
    \arrow["{c'}"', from=2-1, to=2-2]
    \arrow["{r(A')}"', from=2-2, to=2-3]
    \arrow["{\delle_{c'}}"', from=1-1, to=2-1]
    \arrow["f"', from=1-2, to=2-2]
    \arrow["{\delre_c}", from=1-3, to=2-3]
  \end{tikzcd}\]

  And since $r(A) \comp c = c$ and $c' \comp r(A') = c'$, we are finished.

  The other square (i.e., with $c'$ on top) is constructed in an analogous manner.
\end{proof}

@@

    % 3.
    \item We define $\rho^R_{F(c \comp c')}$ as follows.
    % First, we observe that by Lemma \ref{lem:left-rep-by-same-morphism},
    % $F(c \comp c')$ and $Fc \comp Fc'$ are quasi-order-equivalent,
    % as both are left-represented by $Fe_c' \circ Fe_c$.

    First, we claim that $F(c \comp c')$ and $Fc \comp Fc'$ are
    both left-represented by $Fe_{c'} \circ Fe_c$.
    Indeed, we have by part (1) that $e_c' \circ e_c$ left-represents $c \comp c'$,
    and then since $F$ preserves perturbations and squares it follows that
    $F(e_c' \circ e_c) = Fe_{c'} \circ Fe_c$ quasi-left-represents $F(c \comp c')$.
    On the other hand, we also know that $Fe_c$ quasi-left-represents $Fc$ and $Fe_{c'}$
    left-represents $Fc'$, so their composition quasi-left-represents $Fc \comp Fc'$.
    Thus, there is a square $\alpha$ of the form

    % https://q.uiver.app/#q=WzAsNSxbMCwwLCJGQSJdLFsyLDAsIkZBJyciXSxbMCwxLCJGQSJdLFsxLDEsIkZBJyJdLFsyLDEsIkZBJyciXSxbMCwxLCJGKGMgXFxjb21wIGMnKSIsMCx7InN0eWxlIjp7ImJvZHkiOnsibmFtZSI6ImJhcnJlZCJ9LCJoZWFkIjp7Im5hbWUiOiJub25lIn19fV0sWzIsMywiRmMiLDAseyJzdHlsZSI6eyJib2R5Ijp7Im5hbWUiOiJiYXJyZWQifSwiaGVhZCI6eyJuYW1lIjoibm9uZSJ9fX1dLFszLDQsIkZjJyIsMCx7InN0eWxlIjp7ImJvZHkiOnsibmFtZSI6ImJhcnJlZCJ9LCJoZWFkIjp7Im5hbWUiOiJub25lIn19fV0sWzAsMiwiXFxkZWx0YV5sIiwyXSxbMSw0LCJcXGRlbHRhXnIiXSxbOCw5LCJcXGFscGhhIiwxLHsic2hvcnRlbiI6eyJzb3VyY2UiOjIwLCJ0YXJnZXQiOjIwfSwic3R5bGUiOnsiYm9keSI6eyJuYW1lIjoibm9uZSJ9LCJoZWFkIjp7Im5hbWUiOiJub25lIn19fV1d
    \[\begin{tikzcd}[ampersand replacement=\&]
      FA \&\& {FA''} \\
      FA \& {FA'} \& {FA''}
      \arrow["{F(c \comp c')}", "\shortmid"{marking}, no head, from=1-1, to=1-3]
      \arrow["Fc", "\shortmid"{marking}, no head, from=2-1, to=2-2]
      \arrow["{Fc'}", "\shortmid"{marking}, no head, from=2-2, to=2-3]
      \arrow[""{name=0, anchor=center, inner sep=0}, "{\delta^l}"', from=1-1, to=2-1]
      \arrow[""{name=1, anchor=center, inner sep=0}, "{\delta^r}", from=1-3, to=2-3]
      \arrow["\alpha"{description}, draw=none, from=0, to=1]
    \end{tikzcd}\]

    for some perturbations $\delta^l$ and $\delta^r$.
    We define the projection $p_{F(c \comp c')}$ to be $p_{Fc \comp Fc'} \circ \delta^r$.
    We define $\dellp_{F(c \comp c')}$ to be $\dellp_{Fc \comp Fc'} \circ \delta^l$

    Then we can build the square for $\dnr$ by pasting the square $\alpha$ on top of
    the $\dnr$ square for the composition $Fc \comp Fc'$, as shown below:

    % https://q.uiver.app/#q=WzAsNyxbMCwwLCJGQSJdLFsyLDAsIkZBJyciXSxbMCwxLCJGQSJdLFsxLDEsIkZBJyJdLFsyLDEsIkZBJyciXSxbMCwyLCJGQSJdLFsyLDIsIkZBIl0sWzAsMSwiRihjIFxcY29tcCBjJykiLDAseyJzdHlsZSI6eyJib2R5Ijp7Im5hbWUiOiJiYXJyZWQifSwiaGVhZCI6eyJuYW1lIjoibm9uZSJ9fX1dLFsyLDMsIkZjIiwwLHsic3R5bGUiOnsiYm9keSI6eyJuYW1lIjoiYmFycmVkIn0sImhlYWQiOnsibmFtZSI6Im5vbmUifX19XSxbMyw0LCJGYyciLDAseyJzdHlsZSI6eyJib2R5Ijp7Im5hbWUiOiJiYXJyZWQifSwiaGVhZCI6eyJuYW1lIjoibm9uZSJ9fX1dLFs1LDYsInIoRkEpIiwyLHsic3R5bGUiOnsiYm9keSI6eyJuYW1lIjoiYmFycmVkIn0sImhlYWQiOnsibmFtZSI6Im5vbmUifX19XSxbMCwyLCJcXGRlbHRhXmwiLDJdLFsyLDUsIlxcZGVsbHBfe0ZjIFxcY29tcCBGYyd9IiwyXSxbMSw0LCJcXGRlbHRhXnIiXSxbNCw2LCJwX3tGYyBcXGNvbXAgRmMnfSJdLFsxMSwxMywiXFxhbHBoYSIsMyx7InNob3J0ZW4iOnsic291cmNlIjoyMCwidGFyZ2V0IjoyMH0sInN0eWxlIjp7ImJvZHkiOnsibmFtZSI6Im5vbmUifSwiaGVhZCI6eyJuYW1lIjoibm9uZSJ9fX1dLFsxMiwxNCwiXFxkbnJfe0ZjIFxcY29tcCBGYyd9IiwxLHsic2hvcnRlbiI6eyJzb3VyY2UiOjIwLCJ0YXJnZXQiOjIwfSwic3R5bGUiOnsiYm9keSI6eyJuYW1lIjoibm9uZSJ9LCJoZWFkIjp7Im5hbWUiOiJub25lIn19fV1d
    \[\begin{tikzcd}[ampersand replacement=\&]
      FA \&\& {FA''} \\
      FA \& {FA'} \& {FA''} \\
      FA \&\& FA
      \arrow["{F(c \comp c')}", "\shortmid"{marking}, no head, from=1-1, to=1-3]
      \arrow["Fc", "\shortmid"{marking}, no head, from=2-1, to=2-2]
      \arrow["{Fc'}", "\shortmid"{marking}, no head, from=2-2, to=2-3]
      \arrow["{r(FA)}"', "\shortmid"{marking}, no head, from=3-1, to=3-3]
      \arrow[""{name=0, anchor=center, inner sep=0}, "{\delta^l}"', from=1-1, to=2-1]
      \arrow[""{name=1, anchor=center, inner sep=0}, "{\dellp_{Fc \comp Fc'}}"', from=2-1, to=3-1]
      \arrow[""{name=2, anchor=center, inner sep=0}, "{\delta^r}", from=1-3, to=2-3]
      \arrow[""{name=3, anchor=center, inner sep=0}, "{p_{Fc \comp Fc'}}", from=2-3, to=3-3]
      \arrow["\alpha"{marking, allow upside down}, draw=none, from=0, to=2]
      \arrow["{\dnr_{Fc \comp Fc'}}"{description}, draw=none, from=1, to=3]
    \end{tikzcd}\]

    We define $\delrp_{F(c \comp c')}$ to be $\delrp_{Fc \comp Fc'} \circ \delta^r$.
    For $\dnl$, we paste the identity square $\delta^r \ltdyn \delta^r$ on top of
    the $\dnl$ square for the composition $Fc \comp Fc'$, and below that we paste
    the square $\id \ltdyn_{F(c \comp c')}^{Fc \comp Fc'} \id$ which we get from
    the lax property of $F$.
